---
title: "Syllabus"
format:
  html:
    toc: true
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
```

::: {.grid .course-details .course-details-smaller}

::: {.g-col-6}
### Instructor

- {{< fa user >}} &nbsp; [{{< var instructor.name >}}]({{< var instructor.url >}})
- {{< fa university >}} &nbsp; {{< var instructor.office >}}
- {{< fa envelope >}} &nbsp; <a href="mailto:{{< var instructor.email >}}">{{< var instructor.email >}}</a>
- {{< fa calendar-check >}} &nbsp; [Schedule a meeting]({{< var instructor.appointment_url >}})
:::

::: {.g-col-6}
### Course details

- {{< fa calendar >}} &nbsp; {{< var course.days >}}
- {{< fa clock >}} &nbsp; {{< var course.time >}}
- {{< fa location-dot >}} &nbsp; {{< var course.location >}}
- {{< fa calendar-alt >}} &nbsp; {{< var course.dates >}}
:::

:::

## Description

This course explores quantitative research designs to answer questions about public opinion and policy in academia, government, and industry. We will examine how to conduct surveys to understand variation in public opinion or attitudes toward several subjects across the world. We will also examine empirical strategies to generate credible evidence to inform policy decisions in different contexts.

The main learning objective of this course to develop standards to think about the appropriate research design before conducting a study. This is useful to those working with quantitative data either directly or indirectly. We will work towards that goal through a combination of reading, discussion, and hands-on practice.

Students will have the opportunity to practice working with statistical software in preparation to write a pre-analysis plan document that outlines the steps of a future study that addresses a novel question of academic interest or policy/industry relevance.

This course contributes to the Research Methods and/or Analysis requirement for the [Concurrent Certificate in Applied Social Sciences Research](https://academiccalendars.romcmaster.ca/preview_program.php?catoid=47&poid=24300&returnto=9431). As such, it relies heavily on statistics as the main tool to think about research design. Students are expected to have taken at least POLSCI 4NN3 – Statistical Analysis or have equivalent experience with statistics and statistical programming software. I expect us to understand the main findings of a quantitative social science study and work backwards from there to understand the research design choices that researchers made. I also expect, but not assume, some experience with loading, cleaning, and analyzing data.

## Objectives

By the end of the course students should be able to:

- Understand the components of a research design and their properties
- Read and critically evaluate research outputs about public opinion and policy
- Work fluently with statistical programming software and learn new techniques on their own
- Design, evaluate, and implement quantitative studies using the workflow proposed in this course

## Materials

### Reading

The main textbook we will follow is:

- Blair, Graeme, Alexander Coppock, and Macartan Humphreys. 2022. [*Research Design: Declaration, Diagnosis, Redesign*](https://book.declaredesign.org/). Forthcoming at Princeton University Press

You can read the digital copy of the book for free by using the link above. The physical version is not available for purchase yet. The rest of the syllabus refers to this book with the initials RD.

The companion textbook is:

- Grolemund, Garret and Hadley Wickham. 2017. [*R for Data Science*](https://r4ds.had.co.nz/). O’Reilly Media Inc

Once again, you can read the digital copy of this book for free with the link above. You are not required to read this book for this course but consulting it on occasion may help you overcome hurdles while working with statistical software. Since the digital version is constantly updated, I do not recommend buying a physical copy.

We will also read academic papers that discuss or apply the research designs we will cover. Most of these are available through the library’s subscription. See [this link](https://library.mcmaster.ca/services/off-campus-access) for instructions to access library resources while off campus. If not available through the library or elsewhere online, I will upload them to the course website.


### Software

We will use [R](https://www.r-project.org/) and [RStudio](https://posit.co/downloads/) to program research designs and evaluate their properties. The advantage of R is that it is free and open source, meaning that you will be able to apply everything you learn in this course anywhere else. The disadvantage is a somewhat steep learning curve. I believe the investment is worthwhile for anyone working with data or in data-adjacent careers.

The computers in our classroom should have a recent installation of both programs. While I expect us to spend some class meeting time working with R, you will most likely need access to the software outside of the classroom. You are welcome to bring a laptop to class.

You can install R and RStudio in your personal computer. You can use [this link](https://socialsciences.mcmaster.ca/jfox/Courses/R/ICPSR/R-install-instructions.html) for installation instructions on Windows and MacOS (ignore the parts about package building). See [this link](https://levente.littvay.hu/chromebook/) for installation instructions on Chromebooks, which is a bit more involved.

You can also use [Posit Cloud](https://posit.cloud/) to access RStudio from any web browser. A free account should be sufficient for the purposes of this course and has the advantage of letting you access your work across devices.

You should reach out to the instructor if you foresee any challenges with accessing computing resources outside of the classroom.

## Evaluation

See the [Assignments](/assignments/) page for more details.

```{r assignments, include = FALSE}
assignments = tribble(
  ~Assignment, ~Percent,
  "Attendance and participation", "10%",
  "Weekly lab assignments (best 10)", "30% (3% each)",
  "Response papers (best 3)", "30% (10% each)",
  "Final project meetings", "10%",
  "Final project: Pre-analysis plan", "20%"
)
```

<div class="centered-table">

```{r show-assignments, echo = FALSE, results = "asis"}
assignments %>% 
  pandoc.table(justify = "lc")
```

</div>

Grades will be based on the McMaster University scale:

```{r grading, include = FALSE}
grades = tribble(
  ~mark, ~grade, ~mark1, ~grade1,
  "90-100%", "A+", "63-66%", "C",
  "85-90%", "A", "60-62%", "C-",
  "80-84%", "A-", "57-59%", "D+",
  "77-79%", "B+", "53-56%", "D",
  "73-76%", "B", "50-52%", "D-",
  "70-72%", "B-", "0-49%", "F",
  "67-69%", "C+", NA, NA
)
```

<div class="centered-table">

```{r show-grading, echo = FALSE, results = "asis"}
grades %>% 
  pandoc.table(justify = "lclc",
               missing = "",
               col.names = rep(c("Mark", "Grade"), 2))
```

</div>

## Policies

### Submitting assignments

Prompts for assignments will be available in the course website. You should upload assignments via Avenue.

Assignments should use the author-date citation style of the [Chicago Manual of Style](https://www-chicagomanualofstyle-org.libaccess.lib.mcmaster.ca/tools_citationguide/citation-guide-2.html). You do not need to include citations for the weekly lab assignments, but you can do so if you wish.

You can use the templates available in the course website to format your assignments, slight modifications within RStudio are acceptable. If you write assignments outside RStudio, they should be double-spaced, 12 pt font, with 1-inch margins. Assignments do not require a cover sheet. Figures, tables, and bibliography are not part of word counts. You can use [this tool](https://leeper.shinyapps.io/pdfcount/) to count words in PDF documents.

### Late assignments

In this course, assignments are designed to be cumulative; each assignment builds on the last. For this reason, it is important to not fall behind and to complete assignments on time. Assignments turned in within one hour of the due date will only be eligible for 95% of the total value. Assignments turned after one hour but within 24 hours of the due date will be eligible for a maximum grade of B+. Assignments received after 24 hours of the due date will be eligible for a maximum grade of C+. Late assignments will not be accepted after 48 hours after the original due date. 

Use of the MSAF form will automatically move the due date 72 hours, with no other possibility of extension or late submission without additional confirmation of the circumstances by the Faculty advising office. If you use the MSAF form for an assignment, you do not need to email me. Just turn in the assignment within 72 hours via Avenue or to the Political Science office (KTH 527). There is a drop box for after hours.

### Other policies

See the [PDF version of the syllabus](/syllabus.pdf) for McMaster University policies that may apply to this course.



