---
title: "Week 2: Representative Surveys"
date: "January 18, 2023"
format: html
---

```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

We start the topic of public opinion with representative surveys. As the name suggests, a representative survey is administered to a sample of respondents that resemble the demographic characteristics of the underlying population.

The punchline of this week is that conducting a true representative survey via random sampling is near impossible. Instead, most surveys use some form of quota or stratified sampling to ensure that the sample obtained resembles the  population of interest.

The lab assignment for this week will be a guided introduction to R. By the end, we will compare different sampling techniques.

## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD sections 1-4](https://book.declaredesign.org/)

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Schwarz, Norbert. 1999. [“Self-reports: How the questions shape the answers.”](https://dornsife.usc.edu/assets/sites/780/docs/99_ap_schwarz_self-reports.pdf) *American Psychologist* 54 (2): 93-105

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Stephenson, Laura B., Allison Harell, Daniel Rubenson, and Peter John Loewen. 2021. [“Measuring Preferences and Behaviors in the 2019 Canadian Election Study.”](https://doi.org/10.1017/S0008423920001006) *Canadian Journal of Political Science/Revue Canadienne de science politique*


## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/02_slides")
slide_buttons("/slides/02_slides")
```


## Lab

[Intro to R and sampling from populations {{< fa laptop-code >}}](labs/02_lab.Rmd)

::: {.callout-warning}
## Due

Friday, January 20, 5 PM
:::