---
title: "Week 3: Panel Surveys"
date: "January 25, 2023"
format: html
---

```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

Sign up for final project group using [this form {{< fa file-pen >}}](https://forms.gle/6MpV3XfQjpvCd2pS6)

Panel surveys are surveys where the same respondents are asked questions multiple times. Sometimes these repeat the same questions to understand shifts in public opinion and the effect of events in between waves.

As with last week, the takeaway is that panel surveys are expensive and risk missing subjects across waves, this is called panel attrition. With fixed budgets, one may wonder if we are better off just conducting a one-shot survey, also called a cross-section with an increased sample size.

We have no new lab assignment this week. Instead, we will spend some time revisiting the lab from last week. I plan to spend some time comparing one-shot and panel surveys with simulations.

## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD sections 5-7, 15](https://book.declaredesign.org/declaring-designs.html) 

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Esaiasson, Peter and Mikael Persson. 2014. [“Does Studying Political Science Affect Civic Attitudes?: A Panel Comparison of Students of Politics, Law, and Mass Communication.”](https://doi.org/10.1080/15512169.2014.948118) *Journal of Political Science Education* 10 (4): 375-385

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Claes, Ellen, and Marc Hooghe. 2017. [“The Effect of Political Science Education on Political Trust and Interest: Results from a 5-year Panel Study.”](https://doi.org/10.1080/15512169.2016.1171153) *Journal of Political Science Education* 13 (1): 33-45


## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/placeholder")
slide_buttons("/slides/03_slides")
```


## Lab

No new lab! We will revisit the lab from last week.
