---
title: "Survey Experiments"
date: "February 8, 2023"
format: html
---


```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

Some of the techniques we discussed last week to reduce sensitivity bias were **survey experiments**. A survey experiment uses random assignment to administer different pieces of information to respondents, and then compares changes in response to a common outcome. This week, we formalize the **potential outcomes framework** as a way to establish functional relationships in a model. This framework is useful to understand what we can learn in experiments, survey experiments included.

The lab explores the properties of some commonly used survey experimental designs.


## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD section 10](https://book.declaredesign.org/diagnosing-designs.html) 
 
- [{{< fa file-lines >}}]{style="color: #ac1455;"} Tomz, Michael R. and Jessica L. P. Weeks. 2013. [“Public Opinion and the Democratic Peace.”](https://doi.org/10.1017/S0003055413000488) *American Political Science Review* 107 (4): 849-865

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Eggers, Andrew C., Nick Vivyan, and Markus Wagner. 2018. [“Corruption, Accountability, and Gender: Do Female Politicians Face Higher Standards in Public Life?”](https://doi.org/10.1086/694649) *Journal of Politics* 80 (1): 321-326
 

## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/05_slides")
slide_buttons("/slides/05_slides")
```


## Lab

[Survey experimental designs {{< fa laptop-code >}}](labs/05_lab.Rmd)