---
title: "Evidence-Informed Policy"
date: "March 1, 2023"
format: html
---


```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

This week we change topics. We'll start our discussion of public policy with an introduction to the evidence-informed policy movement. This movement emerges from collaborations between governments and academics using cutting-edge statistical techniques to produce more credible evidence to determine what policies work, where, and why.

The lab discusses hypothesis testing, a commonly used approach to quantify statements on whether a policy works.


## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD sections 11-13](https://book.declaredesign.org/redesigning.html)

- Bowers, Jake, and Paul F. Testa. 2019. [“Better Government, Better Science: The Promise of and Challenges Facing the Evidence-Informed Policy Movement.”](https://doi.org/10.1146/annurev-polisci-050517-124041) *Annual Review of Political Science* 22: 521-542

## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/placeholder")
slide_buttons("/slides/08_slides")
```


## Lab

Hypothesis testing {{< fa laptop-code >}} (coming soon!)