---
title: "Field Experiments I"
date: "March 8, 2023"
format: html
---


```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

The next two weeks are about field experiments, also known as randomized controlled trials (RCTs) or A/B testing. These are experiments aimed testing the impact of a policy intervention.

[CONTINUE HERE SO MANY VARIANTS. ALSO ONLY FOCUS ON THE DESIGN FOR THE LONG TERM PROGRAM READING]


## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD section 18](https://book.declaredesign.org/experimental-causal.html)

- Bowers, Jake, and Paul F. Testa. 2019. [“Better Government, Better Science: The Promise of and Challenges Facing the Evidence-Informed Policy Movement.”](https://doi.org/10.1146/annurev-polisci-050517-124041) *Annual Review of Political Science* 22: 521-542

## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/placeholder")
slide_buttons("/slides/09_slides")
```

## Lab

Field experimental designs I {{< fa laptop-code >}} (coming soon!)
