---
title: "Beyond Experimentation"
date: "March 22, 2023"
format: html
---


```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

We use experiments to understand whether a policy would work. We usually care about whether said policy works **outside** of the experimental setting. Experiments alone do not answer that question. Which opens a conversation on the scope and limitations of using experiments to inform policy-making.

The lab illustrates these ideas by trying to compare experiments in different contexts.


## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD section 23](https://book.declaredesign.org/integration.html)

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Deaton, Angus, and Nancy Cartwright. 2018. [“Understanding and misunderstanding randomized controlled trials.”](https://doi.org/10.1016/j.socscimed.2017.12.005) *Social Science & Medicine* 210: 2-21


- [{{< fa file-lines >}}]{style="color: #ac1455;"} Slough, Tara, Daniel Rubenson, Ro’ee Levy, et al. 2021. [“Adoption of community monitoring improves common pool resource management across contexts.”](https://doi.org/10.1073/pnas.2015367118) *Proceedings of the National Academy of Sciences* 118 (29): e2015367118



## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/placeholder")
slide_buttons("/slides/11_slides")
```

## Lab

Limitations of experiments {{< fa laptop-code >}} (coming soon!)
