---
title: "Quasi-experiments II"
date: "April 5, 2023"
format: html
---


```{r setup, echo = FALSE, results = "asis"}
# Slide showing functions
source(here::here("R", "show_slides.R"))
```

## Overview

We continue the conversation on quasi-experiments. This time we focus on difference-in-differences, a research design that relies on repeated observations of units before and after a policy is implemented to make claims about cause and effect. Unlike the previous week, we are not pretending that our data looks as if an experiment was conducted.

The lab discusses difference-in-differences designs.

## Reading

- [{{< fa book >}}]{style="color: #ac1455;"} [RD section 16](https://book.declaredesign.org/observational-causal.html)

- [{{< fa file-lines >}}]{style="color: #ac1455;"} Leninger, Arndt, Marie-Lou Sohnius, Thorsten Faas, Sigrid Roßteutscher, and Armin Schäfer. 2022. [“Temporary Disenfranchisement: Negative Side Effects of Lowering the Voting Age.”](https://doi.org/10.1017/S000305542200034X) *American Political Science Review*

## Slides

::: {.callout-tip}
## Navigation

Click on the frame below and use the arrows to navigate. You can type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) to see useful commands.
:::

```{r show-slides, echo = FALSE, results = "asis"}
show_slide("/slides/placeholder")
slide_buttons("/slides/13_slides")
```

## Lab

Quasi-experimental designs II {{< fa laptop-code >}} (coming soon!)
